Concurrent Priority Queues

### 2014-03-24 presentation/presentation.tex
Add Shavit/Lotan figure
CAS code example
Further work on slides

\usepackage{acronym}
\usepackage[backend=bibtex]{biblatex}
\usepackage{default}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{lmodern}
\usepackage{nameref}
\usepackage{textcomp}

\input{../common/acronyms}
\bibliography{../common/bibliography.bib}

@@

\begin{frame}{Concepts and Definitions}
\framesubtitle{Miscellaneous}

\begin{itemize}
\item \emph{Disjoint-access parallelism}: how well a data structure handles concurrent use by multiple
      threads within disjoint areas
\item Synchronization primitives:
    \begin{itemize}
    \item \ac{CAS}, \ac{FAA}, \ac{FAO}, \ac{TAS}
    \item \ac{DCAS}, \ac{DCSS}
    \end{itemize}
\end{itemize}
\end{frame}

@@

@@

@@

@@

\begin{frame}{\nameref{sec:conclusion}}

>>>>

\usepackage{acronym}
\usepackage[backend=bibtex]{biblatex}
\usepackage{default}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{lmodern}
\usepackage{nameref}
\usepackage{textcomp}

\graphicspath{{figures/}}
\input{../common/acronyms}
\bibliography{../common/bibliography.bib}

@@

\begin{frame}[fragile]{Concepts and Definitions}
\framesubtitle{Miscellaneous}

\begin{itemize}
\item \emph{Disjoint-access parallelism}: how well a data structure handles concurrent use by multiple
      threads within disjoint areas
\item Synchronization primitives:
    \begin{itemize}
    \item \ac{CAS}, \ac{FAA}, \ac{FAO}, \ac{TAS}
    \item \ac{DCAS}, \ac{DCSS}
    \end{itemize}
\end{itemize}

\begin{lstlisting}
bool CAS(T *ptr, T *expected, T value) {
  if (*ptr == *expected) {
    *ptr = value;
    return true;
  } else {
    *expected = *ptr;
    return false;
  }
}
\end{lstlisting}

@@

\begin{figure}
\includegraphics[width=\textwidth]{shavit_lotan}
\caption{The \citeauthor{shavit2000skiplist} \ac{PQ} (Image source: \cite{shavit2000skiplist})}
\end{figure}
\end{frame}

@@

\begin{itemize}
\item Quiescently consistent, but not linearizable
    \begin{itemize}
    \item Slow thread A suspended at deleted key $k$ while in \lstinline|DeleteMin|
    \item Fast thread B first inserts $k-1$, then $k+1$
    \item A wakes up and returns $k+1$
    \item Linearizability would require returning $k-1$
    \end{itemize}

\item Timestamp mechanism: stamp nodes on successful insertion,
      \lstinline|DeleteMin| ignores all stamps earlier than its own starting point
\item Improved scalability, but heavy contention at list head
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:lockfree}}
\framesubtitle{\citeauthor{sundell2003fast}}
% TODO: Way too much content - skip all except Linden?
\begin{itemize}
\item
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:lockfree}}
\framesubtitle{\citeauthor{linden2013skiplist}}

\begin{frame}{\nameref{sec:lockfree}}
\framesubtitle{\citeauthor{linden2013skiplist}}

@@

% Intro relaxed

\begin{frame}{\nameref{sec:relaxed}}
\framesubtitle{\citeauthor{wimmer2013data}}

\begin{frame}{\nameref{sec:relaxed}}
\framesubtitle{\citeauthor{alistarhspraylist}}

% --------------------------------------------------------------------------------------------------
\section{Results} \label{sec:Results}
% --------------------------------------------------------------------------------------------------

\begin{frame}{\nameref{sec:Results}}

\begin{itemize}
\item All benchmarks were performed on Intel Xeon processors
\item Each thread repeatedly chooses uniformly at random between
      \lstinline|Insert| and \lstinline|DeleteMin|
\item Initial priority queue sizes were $2^{15}$ (Gruber), $10^6$ 
      (\citeauthor{alistarhspraylist}), and unspecified (\citeauthor{linden2013skiplist})
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:Results}}
\framesubtitle{Gruber}

\begin{figure}
% TODO: Sweave
\caption{Throughput benchmark results.}
\end{figure}
\end{frame}

\begin{frame}{\nameref{sec:Results}}
\framesubtitle{\citeauthor{linden2013skiplist}}

\begin{figure}
\includegraphics[width=0.6 \textwidth]{results_linden}
\caption{Throughput benchmark results.
\emph{New}: \citeauthor{linden2013skiplist},
\emph{HaS}: \citeauthor{shavit2000skiplist},
\emph{SaT}: \citeauthor{sundell2003fast}. 
(Image source: \cite{linden2013skiplist})}
\end{figure}
\end{frame}

\begin{frame}{\nameref{sec:Results}}
\framesubtitle{\citeauthor{alistarhspraylist}}

\begin{figure}
\includegraphics[width=\textwidth]{results_spraylist}
\caption{Throughput benchmark results. 
(Image source: \cite{alistarhspraylist})}
\end{figure}
\end{frame}

@@

\begin{frame}{\nameref{sec:conclusion}}
\begin{itemize}
\item Parallelizing \acp{PQ} is hard
\item SkipLists currently dominate practical implementations
\item Main limiting factor are list head accesses in \lstinline|DeleteMin|
\item \citeauthor{linden2013skiplist} \ac{PQ} is state of the art in strict
      semantics
\item Much potential remains in relaxed \acp{PQ} and data structures in general
      \textrightarrow ~ future research
\item A standalone version of the \citeauthor{wimmer2013data} \ac{PQ} is planned
      and will allow direct comparison against related data structures
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:conclusion}}
\centering
Questions?
\end{frame}
